---
title: "Introduction to the New Template"
teaching: 10
exercises: 2
---

:::::::::::: questions

 - How do I get started?

::::::::::::::::::::::

::::::::::: objectives

- Create new lesson from scratch
- Identify the main command to preview the template
- Understand the basic structure of the new template

::::::::::::::::::::::


## Quickstart: Create a New Lesson

Let's say you have a set of R Markdown files that you used for a class website
that you want to convert into a Carpentries Lesson. To go from zero to a new 
lesson website that auto-renders R Markdown documents to a functional website is
three steps with `{sandpaper}`:

1. Create a site
2. Push to GitHub
3. Add your files

That's it. After that, if you know how to write in Markdown, you are good to go.

:::::::::::::::::: callout

### Takeaway message

Contributors should only be expected to know basic markdown and *very* minimal
yaml syntax in order to work on lessons.

:::::::::::::::::::::::::::

::::::::::::::::: challenge

### Try it yourself!

Follow these steps to create a brand new lesson and push it to GitHub

1. Follow the [setup instructions](setup.html)
2. Open RStudio (or your preferred interface to R)
3. Use the following code:

```r
library("usethis")
library("sandpaper")

# Create a brand new lesson on your desktop called "bouyant-barnacle"
create_lesson("~/Desktop/bouyant-barnacle")
# Push the lesson to GitHub
use_github()
```


::::::::::: solution

If everything went correctly, you will now have RStudio open to your new project
called "bouyant-barnacle" on your Desktop and you will have a brand new
repository on your GitHub account called "bouyant-barnacle". 

In a few minutes, the GitHub workflows would have validated your lesson and 
created deployment branches. You can track the progress at 
`https://github.com/<USERNAME>/bouyant-barnacle/actions/`. Once you have a green
check mark, you can [set up GitHub Pages](https://docs.github.com/en/github/working-with-github-pages/configuring-a-publishing-source-for-your-github-pages-site) by 
choosing `gh-pages` from the dropdown menu.

> **Be patient!** GitHub needs to start up a new virtual machine the first time
> you use this, so it may take anywhere from 6 to 15 minutes for the workflow
> to run.

::::::::::::::::::::
::::::::::::::::::::::::

## Previewing Your New Lesson

After you created your lesson, you will want to preview it locally. There is one
command to do that in [{sandpaper}]:

```{r, eval = FALSE}
sandpaper::build_lesson()
```

:::::::::::::: callout

### DID YOU KNOW? Keyboard Shortcuts are Available

If you are using RStudio, did you know you can use keyboard shortcuts to render
the lesson as you are working on the episodes?

Render and preview the whole lesson
:    <kbd>ctrl + shift + B</kbd>

Render and preview an episode
:    <kbd>ctrl + shift + K</kbd>
::::::::::::::::::::::

The first time you run this function, you might see A LOT of output on your
screen and then your browser will open the preview. If you run the command 
again, you will see much less output. If you like to would like to know how
everything works under the hood, you can check out the [{sandpaper} generator
chapter](sandpaper.html).

## Folder Structure

:::::::::::::::: callout

### :construction: This May Change :construction:

The exact folder structure still has the possibility to change based on user
testing for the front-end of the lesson website.

::::::::::::::::::::::::

The template folder structure will contain markdown files arranged so that they
match what we expect the menubar for the lesson should be. All folders and files
with an arrow `<-` are places in the lesson template you will be modifying:

```
|-- .gitignore         #  | Ignore everything in the site/ folder
|-- .github/           #  | Configuration for deployment
|-- episodes/          # <- PUT YOUR EPISODE MARKDOWN FILES IN THIS FOLDER
|-- instructors/       # <- Information for Instructors (e.g. guide.md)
|-- learners/          # <- Information for Learners (e.g. reference.md and setup.md)
|-- profiles/          # <- Learner and/or Instructor Profiles (new)
|-- site/              #  | This is a "scratch" folder ignored by git and is where the rendered markdown files and static site will live
|-- config.yaml        # <- Use this to configure lesson metadata
|-- index.md           # <- The landing page of your site
|-- CONTRIBUTING.md    #  | Carpentries Rules for Contributions (REQUIRED)
|-- CODE_OF_CONDUCT.md #  | Carpentries Code of Conduct (REQUIRED)
|-- LICENSE.md         #  | Carpentries Licenses (REQUIRED)
`-- README.md          # <- Introduces folks how to use this lesson and where they can find more information.
```

This folder structure is heavily opinionated towards achieving our goals of
creating a lesson template that is fit for the purpose of delivering lesson
content for not only Carpentries instructors, but also for learners and
educators who are browsing the content after a workshop. It is not designed to
be a blog or commerce website. 

## Tools

As described in [the setup document](setup.html), the lesson template now only
requires R and [pandoc] to be installed. The tooling from the current lesson
template has been split up into three R packages:

1. [{varnish}] contains the HTML, CSS, and JavaScript elements
1. [{pegboard}] is a validator for the markdown documents
1. [{sandpaper}] is the engine that puts everything together. 

::::::::::::: keypoints

- Lessons can be created with `create_lesson()`
- Preview lessons with `build_lesson()`
- To edit a lesson, you only need to know Markdown and/or R Markdown
- The folder structure is designed with maintainers in mind
- The toolchain is designed to be modular.

:::::::::::::::::::::::


<!-- Please do not delete anything below this line -->

[{varnish}]: template.html
[{pegboard}]: validator.html
[{sandpaper}]: engine.html
[The Carpentries Incubator]: https://carpentries.org/community-lessons/
[^history]: An incomplete history is noted in the 9.1.0 release: https://github.com/carpentries/styles/releases/tag/v9.1.0

```{r links, child = sandpaper::template_links()}
```
